---
title: "econ725 project Q3"
author: "Demi Han, Ruyuan Mei"
date: "11/19/2020"
output: pdf_document
---


######Sugandh###### 
```{r message=FALSE, warning=FALSE}
require(tidyverse)
require(ggplot2)
require(data.table)
require(plyr)
require(dplyr)
require(knitr)
require(foreign)
require(ggcorrplot)
require(corrplot)
require(caret)
require(gridExtra)
require(scales)
require(Rmisc)
require(ggrepel)
require(randomForest)
require(glmnet)
require(psych)
require(xgboost)
require(ggthemes)
```


```{r message=FALSE, warning=FALSE}
#loading the dataset
df <- read.dta("~/Desktop/ebaydatafinal.dta")
```

```{r}
#summary for the highest bid
summary(df$biddy1)
```

There are 22522 null values in highest bid variable. This column will be revenue as its the amount the seller gets when he sells the item. one thing to notice here is that the maximum bid in dataset is 1780400.

```{r}
#keeping only items which have been sold
#df <- df[df$sell == 1 ,]
```

I did this considering that if the item isn't sold , then there is no revenue for the seller.

Checking if we have null values now for highest bid 

```{r}
summary(df$biddy1)
```



## Data Cleaning 

# Formatting Dates: 

The columns start date and end date 

```{r}
head(df$startdate)
```

```{r message=FALSE, warning=FALSE}
library("lubridate")
#converting strings into date format 
df$startdate <- parse_date_time(df$startdate, orders="mdy HMS")
df$enddate <- parse_date_time(df$enddate, orders="mdy HMS")
```


```{r}
#extracting months from dates 
df$months <-  month(df$startdate)
df$days <- day(df$startdate)
df$monthe <-  month(df$enddate)
df$daye <- day(df$enddate)
```

```{r}
#converting long dates to short dates and converting them to mm-dd-yy format 
df$startdate <- date(df$startdate)
df$startdate <- format(df$startdate, "%m-%d-%y")
df$enddate <- date(df$enddate)
df$enddate <- format(df$enddate, "%m-%d-%y")
```


The most importent numeric variables 

```{r}
numericVars <- which(sapply(df, is.numeric)) #index vector numeric variables
numericVarNames <- names(numericVars) #saving names vector for use later on
cat('There are', length(numericVars), 'numeric variables')
```

```{r}
df_numVar <- df[, numericVars]
#correlation of all numeric variables
cor_numVar <- cor(df_numVar, use="pairwise.complete.obs")
```

```{r}
#sort on decreasing correlations with highest bid
cor_sorted <- as.matrix(sort(cor_numVar[,'biddy1'], decreasing = TRUE))
```



Lets see which variables are positively correlated with highest bid 

```{r}
head(cor_sorted ,50)
```

From above we can see that biddy2 , bookvalue , startbid , warranty , options , phone, logsize , loghtml , logtext , numbids , featured , descriptionsize , dealer , length , inspection , photos , logage are highly correlated with highest bid .



Now , lets see which variables are negatively correlated with highest bid 
```{r}
tail(cor_sorted ,50)
```


From above we can see that logmiles , rust , dent , crack , age , broken , problem are negatively correlated with highest bid . 


## Missing data , label encoding and Factorizing variables 



```{r}
#which columns have missing values 
NAcol <- which(colSums(is.na(df)) > 0)
NAcol
```

```{r}
cat('There are', length(NAcol), 'columns with missing values')
```
bookvalue has 19 missing values and photos has 29 missing values and biddy5 has  45 missing values , for now I am just dropping these missing values and we will think about imputingf them in future . 

```{r message=FALSE}
#deleting missing values 
df=df[!is.na(df$bookvalue),]
df=df[!is.na(df$photos),]
df=df[!is.na(df$biddy5),]
```


Now lets try imputing age and logmiles variables. I am imputing these variables with the median 

```{r message=FALSE, warning=FALSE}
library(Hmisc)
df$age<-impute(df$age, median)
df$logmiles<-impute(df$logmiles, median)
```



## Label Encoding / factorizing the charachter variables 

```{r}
Charcol <- names(df[,sapply(df, is.character)])
Charcol
```

```{r}
cat('There are', length(Charcol), 'remaining columns with character values')
```
First lets consider variables maker , interior and exterior . They all are factor variables .

```{r}
df$maker <- as.factor(df$maker)
table(df$maker)
```

```{r}
df$interior <- as.factor(df$interior)
table(df$interior)
```

```{r}
df$exterior <- as.factor(df$exterior)
table(df$exterior)
```


# dealing with date variables 

```{r}
df$membersince <- parse_date_time(df$membersince, orders="mdy")
df$monthm <-month(df$membersince)
df$daym <- day(df$membersince)
df$membersince <- date(df$membersince)
df$membersince <- format(df$membersince, "%m-%d-%y")
```

```{r}
df$months <- as.factor(df$months)
df$days <- as.factor(df$days)
df$monthe <- as.factor(df$monthe)
df$daye <- as.factor(df$daye)
df$monthm <- as.factor(df$monthm)
df$daym <- as.factor(df$daym)
```


## Correlations

```{r}
#keeping only required columns 
df<-df[, c("biddy1" , "biddy2" , "biddy3" ,"biddy4", "biddy5" ,"bookvalue", "photos",  "startbid" , "warranty" , "options" , "phone", "logsize" , "loghtml" , "logtext" , "numbids" , "featured" , "descriptionsize" , "dealer" , "length" , "inspection" , "photos" , "age" , "logmiles" , "rust" , "dent" , "crack" , "age" , "broken","months", "monthm", "monthe", "days" , "daye" , "daym", "maker", "interior" , "exterior" , "enddate" , "startdate")]
```


## Finding variable importance with Random forest 

```{r}
set.seed(2020)
quick_RF <- randomForest(x=df[1:13374,-36], y= df$biddy1[1:13374], ntree=100,importance=TRUE)
imp_RF <- importance(quick_RF)
imp_DF <- data.frame(Variables = row.names(imp_RF), MSE = imp_RF[,1])
imp_DF <- imp_DF[order(imp_DF$MSE, decreasing = TRUE),]
ggplot(imp_DF[1:20,], aes(x=reorder(Variables, MSE), y=MSE, fill=MSE)) + geom_bar(stat = 'identity') + labs(x = 'Variables', y= '% increase MSE if variable is randomly permuted') + coord_flip() + theme(legend.position="none")
```



Lets draw some graphs associated with the highest bid/ revenue. first lets see the density of the biddy1 

```{r}
p2 <-ggplot(data=df[df$biddy1 < 50000,], aes(x= biddy1))+
  geom_histogram(fill="blue", binwidth = 1500)+
  ggtitle('Highest bid density ') + theme_stata()
p2
```

Lets look at the relationship of this biddy1/revenue with number of bidders 
```{r}
p2 <-ggplot(data= df[df$biddy1 < 50000,], aes(x = numbids,y= biddy1))+
  geom_point()
p2
```




######Demi Han, Ruyuan Mei######


```
``{r failed}
#split biddy1 by month
df1<-df[, c("biddy1" ,"months", "monthe", "days" , "daye" , "enddate" , "startdate")]
sp<-split(df1,df1[,c("monthe")],drop=TRUE)
sp1<-data.table(sp)
result1<-lapply(sp1,FUN=function(x) sum(x$AMOUNT))
result2<-lapply(sp1,FUN=function(x) mean(x$AMOUNT))
result<-cbind(result1,result2)
```

```
``{r failed}
#split biddy1 by month
df1<-df[, c("biddy1" ,"months" , "monthe", "days" , "daye" , "enddate" , "startdate")]


g <- split(df1,df1$monthe,)
g
df2<-data.table(g)

g1 <- lapply(df2,mean)
result1<-lapply(df2,FUN=function(x) sum(df2$biddy1))
df3 <- data.table(result1)


sp<-split(df1,df1[,c("monthe","biddy1")],drop=TRUE)
result1<-lapply(sp,FUN=function(x) sum(x$AMOUNT))
result2<-lapply(sp,FUN=function(x) mean(x$AMOUNT))
result<-cbind(result1,result2)
```


```{r}
#split biddy1 by month
df1<-df[, c("biddy1" ,"months" , "monthe", "days" , "daye" , "enddate" , "startdate")]
dfmonth2 <- df1[df1$monthe == 2,]
meanmonth2 <- mean(dfmonth2$biddy1)
summonth2 <- sum(dfmonth2$biddy1)
meanmonth2
summonth2

dfmonth3 <- df1[df1$monthe == 3,]
meanmonth3 <- mean(dfmonth3$biddy1)
summonth3 <- sum(dfmonth3$biddy1)
meanmonth3
summonth3

dfmonth4 <- df1[df1$monthe == 4,]
meanmonth4 <- mean(dfmonth4$biddy1)
summonth4 <- sum(dfmonth4$biddy1)
meanmonth4
summonth4

dfmonth5 <- df1[df1$monthe == 5,]
meanmonth5 <- mean(dfmonth5$biddy1)
summonth5 <- sum(dfmonth5$biddy1)
meanmonth5
summonth5

dfmonth6 <- df1[df1$monthe == 6,]
meanmonth6 <- mean(dfmonth6$biddy1)
summonth6 <- sum(dfmonth6$biddy1)
meanmonth6
summonth6

dfmonth7 <- df1[df1$monthe == 7,]
meanmonth7 <- mean(dfmonth7$biddy1)
summonth7 <- sum(dfmonth7$biddy1)
meanmonth7
summonth7

dfmonth8 <- df1[df1$monthe == 8,]
meanmonth8 <- mean(dfmonth8$biddy1)
summonth8 <- sum(dfmonth8$biddy1)
meanmonth8
summonth8

dfmonth9 <- df1[df1$monthe == 9,]
meanmonth9 <- mean(dfmonth9$biddy1)
summonth9 <- sum(dfmonth9$biddy1)
meanmonth9
summonth9

dfmonth10 <- df1[df1$monthe == 10,]
meanmonth10 <- mean(dfmonth10$biddy1)
summonth10 <- sum(dfmonth10$biddy1)
meanmonth10
summonth10

#summary monthly mean
monthe <- c(2,3,4,5,6,7,8,9,10)
mean <- c(meanmonth2,meanmonth3,meanmonth4,meanmonth5,meanmonth6,meanmonth7,meanmonth8,meanmonth9,meanmonth10)
#summary monthly sum
sum <- c(summonth2,summonth3,summonth4,summonth5,summonth6,summonth7,summonth8,summonth9,summonth10)

monthly <- data.frame(monthe,mean,sum)
monthly
```

```{r}
#graphically biddy1's mean by month
monthly <- tibble(
  month = c("2","3","4","5","6","7","8","9","10"),
  mean = c(monthly$mean)
)
knitr::kable(monthly)

p3 <- ggplot(data = monthly, mapping = aes(
  x = fct_reorder(month, desc(mean)),
  y = mean ))

p3 + geom_col(fill = "lightblue") +
  geom_text(mapping = aes(
    y = mean / 2, label = paste(mean))) +
  scale_y_continuous(breaks = NULL) +
  coord_flip() + 
  labs(x = "month", 
       y = "mean")
```



```{r}
#summary by days
df1$daye<- as.numeric(df1$daye)
dfmonth_b <- df1[df1$daye <= 10,]
meanmonth_b <- mean(dfmonth_b$biddy1)
summonth_b <- sum(dfmonth_b$biddy1)
meanmonth_b
summonth_b

dfmonth_m <- df1[df1$daye >=11 & df1$daye <= 20,]
meanmonth_m <- mean(dfmonth_m$biddy1)
summonth_m <- sum(dfmonth_m$biddy1)
meanmonth_m
summonth_m

dfmonth_e <- df1[df1$daye >= 21 & df1$daye <= 31,]
meanmonth_e <- mean(dfmonth_e$biddy1)
summonth_e <- sum(dfmonth_e$biddy1)
meanmonth_e
summonth_e

#summary mean
period <- c("Beginning of month","Middle of month","Ending of month")
meandays <- c(meanmonth_b,meanmonth_m,meanmonth_e)
#summary sum
sumdays <- c(summonth_b,summonth_m,summonth_e)

daily <- data.frame(period,meandays,sumdays)
daily
daily <- data.table(daily)
```


```{r}
#graphically biddy1's mean by days
daily <- tibble(
  period = c(daily$period),
  meandays = c(daily$meandays)
)
knitr::kable(daily)

p4 <- ggplot(data = daily, mapping = aes(
  x = fct_reorder(period, desc(meandays)),
  y = meandays ))

p4 + geom_col(fill = "orange", width = 0.4) +
  geom_text(mapping = aes(
    y = meandays / 2, label = paste(meandays))) +
  scale_y_continuous(breaks = NULL) +
  coord_flip() +
  labs(x = "period", 
       y = "mean")
```
#In conclusion, as buyers, we could get a better price in the endding of the month, and avoid buying a car in March to June.


```{r message=FALSE, warning=FALSE}
#weekdays
df <- read.dta("~/Desktop/ebaydatafinal.dta")
df <- df[df$sell == 1 ,]
library("lubridate")
#converting strings into date format 
df$startdate <- parse_date_time(df$startdate, orders="mdy HMS")
df$enddate <- parse_date_time(df$enddate, orders="mdy HMS")
df$wdays <-  wday(df$enddate)

df4 <-df[, c("biddy1" ,  "enddate" , "wdays")]
```

```{r}
#summary by weekdays
df4$wdays <- as.numeric(df4$wdays)

dfweekday <- df4[df4$wdays <= 5,]
meanweekday <- mean(dfweekday$biddy1)
meanweekday

dfweekend <- df4[df4$wdays > 5,]
meanweekend <- mean(dfweekend$biddy1)
meanweekend
```

```{r}
#graphically weekdays vs. weekend
wdays <- c("weekday","weekend")
meanwdays <- c(meanweekday,meanweekend)
weekly <- data.frame(wdays,meanwdays)
weekly
weekly <- data.table(weekly)
```

```{r}
#graphically biddy1's mean (weekdays vs. weekend)
weekly <- tibble(
  wdays = c(weekly$wdays),
  meanwdays = c(weekly$meanwdays)
)
knitr::kable(weekly)

p5 <- ggplot(data = weekly, mapping = aes(
  x = wdays,
  y = meanwdays ))
p5 + geom_col(fill = "orange",width = 0.4) +
  geom_text(mapping = aes(
    y = meanwdays / 2, label = paste(meanwdays))) +
  scale_y_continuous(breaks = NULL) +
  coord_flip() +
  labs(x = "wdays", 
       y = "mean")

```
